# 1. Show a graphical overview of the data and show summaries of the variables in the data. Describe and interpret the outputs, commenting on the distributions of the variables and the relationships between them. (0-3 points)

```{r}
human <- read.table("http://s3.amazonaws.com/assets.datacamp.com/production/course_2218/datasets/human1.txt", sep  =",", header = T, row.names = "Country")
names(human)
str(human)
head(human)
summary(human)
```

The data was collected by the UN to measure Human Development Index (HDI) around the world. The variables refer to, among others, life expectancy, education, gender inequality, etc.



#2. Perform principal component analysis (PCA) on the not standardized human data. Show the variability captured by the principal components. Draw a biplot displaying the observations by the first two principal components (PC1 coordinate in x-axis, PC2 coordinate in y-axis), along with arrows representing the original variables. (0-2 points)


```{r}
summary(human)
pca_human <- prcomp(human)
```

