# Longitudinal Data # TASK 2 BPRS

#Read the data
```{r}
BPRS <- read.csv(file = "~/IODS-project/data/BPRS.csv", header = TRUE, row.names = 1)
BPRSL <- read.csv(file = "~/IODS-project/data/BPRSL.csv", header = TRUE, row.names = 1)
library(dplyr)
library(tidyr)
library(ggplot2)
```

#Factor variables & glimpse
```{r}
BPRS$treatment <- factor(BPRS$treatment)
BPRS$subject <- factor(BPRS$subject)
glimpse(BPRS)
```

```{r}
BPRSL <-  BPRS %>% gather(key = weeks, value = bprs, -treatment, -subject)
BPRSL <-  BPRSL %>% mutate(week = as.integer(substr(weeks,5,5)))
glimpse(BPRSL)

```


#Linear mixed effects models 

```{r}
dim(BPRSL)
```

#Standardizing the data

```{r}
BPRSL <- BPRSL %>%
  group_by(week) %>%
  mutate(stdbprs = (bprs - mean(bprs))/sd(bprs) ) %>%
  ungroup()
```

#Plot
```{r}
ggplot(BPRSL, aes(x = week, y = bprs, linetype = subject)) +
  geom_line() +
  scale_linetype_manual(values = rep(1:10, times=4)) +
  facet_grid(. ~ treatment, labeller = label_both) +
  theme(legend.position = "none") + 
  scale_y_continuous(limits = c(min(BPRSL$bprs), max(BPRSL$bprs)))
```

#Plot
```{r}
ggplot(BPRSL, aes(x = week, y = bprs, group = subject)) +
  geom_line(aes(linetype = subject)) +
  scale_x_continuous(name = "Week", breaks = seq(0, 60, 10)) +
  scale_y_continuous(name = "bprs") +
  theme(legend.position = "top")

```


```{r}
ggplot(BPRSL, aes(x = week, y = bprs, linetype = subject)) +
  geom_line() +
  scale_linetype_manual(values = rep(1:10, times=4)) +
  facet_grid(. ~treatment, labeller = label_both) +
  theme(legend.position = "none") + 
  scale_y_continuous(limits = c(min(BPRSL$bprs), max(BPRSL$bprs)))
```

#Multiple linear regression model
```{r}
BPRS_reg <- lm(bprs ~ week + treatment, data = BPRSL)
summary(BPRS_reg)
```

#The Random Intercept Model

```{r}
library(lme4)
BPRS_ref <- lmer(bprs ~ week + treatment + (1 | subject), data = BPRSL, REML = FALSE)
summary(BPRS_ref)
```

#Slippery slopes: Random Intercept and Random Slope Model
```{r}
BPRS_ref1 <- lmer(bprs ~ week + treatment + (week | subject), data = BPRSL, REML = FALSE)
summary(BPRS_ref1)
```


#Anova
```{r}
anova(BPRS_ref1, BPRS_ref)
```


#Random Intercept and Random Slope Model with interaction

```{r}
BPRS_ref2 <- lmer(bprs ~ week * treatment + (week | subject), data = BPRSL, REML = FALSE)
```

#Print summary

```{r}
summary(BPRS_ref2)
```

# perform an ANOVA test on the two models

```{r}
anova(BPRS_ref2, BPRS_ref1)
```

#Draw the plot
```{r}
library(ggplot2)
library(dplyr) 
library(tidyr)
library(lme4)
ggplot(BPRSL, aes(x = week, y = bprs, group = treatment)) +
  geom_line(aes(linetype = treatment)) +
  scale_x_continuous(name = "week", breaks = seq(0, 60, 20)) +
  scale_y_continuous(name = "bprs") +
  theme(legend.position = "top")
```

# Create a vector of the fitted values
```{r}
Fitted <- fitted(BPRS_ref2)
```
# Create a new column fitted to BPRSL

```{r}
BPRSL <- BPRSL %>%
  mutate(Fitted)
```

#Draw plot
```{r}
ggplot(BPRSL, aes(x = week, y = Fitted, group = treatment)) +
  geom_line(aes(linetype = treatment)) +
  scale_x_continuous(name = "Week", breaks = seq(0, 60, 20)) +
  scale_y_continuous(name = "Bprs") +
  theme(legend.position = "top")

```



