
#Longitudinal Data # TASK 1

# TASK 1:Implement the analysis of Chapter 8 of MABS using the RATS data. (0-7 points: 0-4 points for graphs or analysis results + 0-3 points for their interpretations)

```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
RATS <- read.csv(file = "~/IODS-project/data/RATS.csv", header = TRUE, row.names = 1)
RATSL <-read.csv(file = "~/IODS-project/data/RATSL.csv", header = TRUE, row.names = 1)
dim(RATSL)
RATS$ID <- factor(RATS$ID)
RATS$Group <- factor(RATS$Group)
RATSL$ID <- factor(RATSL$ID)
RATSL$Group <- factor(RATSL$Group)
glimpse(RATS)
ggplot(RATSL, aes(x = Time, y = Weight, group = ID)) +
  geom_line(aes(linetype = Group)) +
  scale_x_continuous(name = "Time (days)", breaks = seq(0, 60, 10)) +
  scale_y_continuous(name = "Weight (grams)") +
  theme(legend.position = "top")
```

#Tracking. The graph shows the weight changes of rats allocated in three observed groups during the period of over 60 days of study. The Group 1, with the lowest weight at the starting point demonstrates the most modest weight fluctuation when compared with the two other groups, and remain as the one characterized with the lowest weight values. The groups 2 and 3, in turn, show comparatively quite a significant weight gain, even up to around 25% of the starting weight in at least one case. To sum up, the groups 2 and 3 demonstrate tendency towards weight gain, while the group 1 is characterized by almost constant weight throughout the study. 

#More detailed view of the individual rats, divided into groups:

```{r}
library(ggplot2)
ggplot(RATSL, aes(x = Time, y = Weight, linetype = ID)) +
  geom_line() +
  scale_linetype_manual(values = rep(1:10, times=4)) +
  facet_grid(. ~ Group, labeller = label_both) +
  theme(legend.position = "none") + 
  scale_y_continuous(limits = c(min(RATSL$Weight), max(RATSL$Weight)))
```

Here we can better see the individual rats and their progress during the study.  

#Standardizing the variables
```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
RATSL <- RATSL %>%
  group_by(Time) %>%
  mutate(stdbprs = (Weight - mean(Weight))/sd(Weight) ) %>%
  ungroup()
glimpse(RATSL)
```

# Plot again using the standardised data

```{r}
ggplot(RATSL, aes(x = Time, y = Weight, linetype = ID)) +
  geom_line() +
  scale_linetype_manual(values = rep(1:10, times=4)) +
  facet_grid(. ~ Group, labeller = label_both) +
  scale_y_continuous(name = "Rats - standardized")
```

# The summary graph:

```{r}
n <- RATSL$Time %>% unique() %>% length()
RATSS <- RATSL %>%
  group_by(Group, WD) %>%
  summarise( mean = mean(Weight), se = sd(Weight)/sqrt(n) ) %>%
  ungroup()
glimpse(RATSS)
```

#Plotting the mean profiles

```{r}
ggplot(RATSS, aes(x = WD, y = mean, linetype = "ID", shape = "ID")) +
  geom_line() +
  scale_linetype_manual(values = c(1,2)) +
  geom_point(size=3) +
  scale_shape_manual(values = c(1,2)) +
  geom_errorbar(aes(ymin = mean - se, ymax = mean + se, linetype="1"), width=0.3) +
  theme(legend.position = c(0.8,0.8)) +
  scale_y_continuous(name = "mean(Weight) +/- se(Weight)")
```
# The plot above shows the mean response proﬁles for the individual rats. This kind of summary measure analysis of longitudinal data takes into account singular, repeated  measurements of the individuals taken during the study and expresses them in form of a single value.  

```{r}
RATSL8S <- RATSL %>%
  filter(Time > 0) %>%
  group_by(Group, ID) %>%
  summarise( mean=mean(Weight) ) %>%
  ungroup()
glimpse(RATSL8S)
```

#The summary measure approach - the way to reveal an outlier, which may bias the conclusions:
```{r}
ggplot(RATSL8S, aes(x = Group, y = mean)) +
  geom_boxplot() +
  stat_summary(fun.y = "mean", geom = "point", shape=23, size=4, fill = "pink") +
  scale_y_continuous(name = "mean(Weight), weeks 1-8")
```

# Filtering the outlier & drawing the plot:
```{r}
RATSL8S1 <- RATSL8S %>%
  filter(mean > 550)
ggplot(RATSL8S, aes(x = Group, y = mean)) +
  geom_boxplot() +
  stat_summary(fun.y = "mean", geom = "point", shape=23, size=4, fill = "pink") +
  scale_y_continuous(name = "mean(Weight), weeks 1-8")
```

#T-test is performed using: t.test(mean ~ Group, data = RATSL8S1, var.equal = TRUE) 

#Anova: RATSL8S2 <- RATSL8S %>%, mutate(baseline = RATS$week0), fit <- lm(mean ~ baseline + Group, data = RATSL8S2), anova(fit)
