# Beata, 20.11.2019

```{r}
library(MASS)
data("Boston")
str(Boston)
summary(Boston)
pairs(Boston)
```

Description: The collected data is related to the housing values in suburbs of Boston, MA, US. The data frame has 506 rows and 14 columns. Such factors like crime rate, proportion of non-retail business, avarage number of rooms, and so on, are taken into account. 


```{r}
library(MASS)
data("Boston")
library(corrplot)
library(tidyr)
cor_matrix<-cor(Boston)  %>% round(digits=2)
cor_matrix
corrplot(cor_matrix, method="circle", type="upper", cl.pos="b", tl.pos="d", tl.cex = 0.6)

```

```{r}
library(MASS)
data(Boston)
boston_scaled <- scale(Boston)
boston_scaled <- as.data.frame(boston_scaled)
summary(boston_scaled$crim)
bins <- quantile(boston_scaled$crim)
bins
crime <- cut(boston_scaled$crim, breaks = bins, include.lowest = TRUE, labels = c("low", "med_low", "med_high", "high"))
table(crime)
boston_scaled <- dplyr::select(boston_scaled, -crim)
boston_scaled <- data.frame(boston_scaled, crime)

```

```{r}
n <- nrow(boston_scaled)
ind <- sample(n,  size = n * 0.8)
train <- boston_scaled[ind,]
test <- boston_scaled[-ind,]
correct_classes <- test$crime
test <- dplyr::select(test, -crime)

```

```{r}
library(MASS)
#train <- boston_scaled[ind,]
#data(train)
lda.fit <- lda(crime ~ ., data = train)
lda.fit
lda.arrows <- function(x, myscale = 1, arrow_heads = 0.1, color = "orange", tex = 0.75, choices = c(1,2)){
  heads <- coef(x)
  arrows(x0 = 0, y0 = 0, 
         x1 = myscale * heads[,choices[1]], 
         y1 = myscale * heads[,choices[2]], col=color, length = arrow_heads)
  text(myscale * heads[,choices], labels = row.names(heads), 
       cex = tex, col=color, pos=3)
}

classes <- as.numeric(train$crime)
plot(lda.fit, dimen = 2, col = classes, pch = classes)
lda.arrows(lda.fit, myscale = 1)

```

```{r}
lda.pred <- predict(lda.fit, newdata = test)
table(correct = correct_classes, predicted = lda.pred$class)
```

```{r}
library(MASS)
data(Boston)
scale(Boston)
dist_eu <- dist(Boston)
summary(dist_eu)
dist_man <- dist(Boston, method = 'manhattan')
summary(dist_man)

```


BONUS: Bonus: Perform k-means on the original Boston data with some reasonable number of clusters (> 2). Number of clusters: 4. 
```{r}
data(Boston)
scale(Boston)
km <-kmeans(Boston, centers = 4)
pairs(Boston, col = km$cluster)
```

```{r}
data(Boston)
set.seed(123)
k_max <- 10
twcss <- sapply(1:k_max, function(k){kmeans(Boston, k)$tot.withinss})
library(ggplot2)
qplot(x = 1:k_max, y = twcss, geom = 'line')
km <-kmeans(Boston, centers = 2)
pairs(Boston, col = km$cluster)

```

