---
output:
  html_document: default
  pdf_document: default
---
#*TEA TIME*

#Load the tea dataset from the package Factominer. Explore the data briefly: look at the structure and the dimensions of the data and visualize it. Then do Multiple Correspondence Analysis on the tea data (or to a certain columns of the data, it’s up to you). Interpret the results of the MCA and draw at least the variable biplot of the analysis. You can also explore other plotting options for MCA. Comment on the output of the plots. (0-4 points)

```{r}
library(FactoMineR)
library(ggplot2)
library(dplyr)
library(tidyr)
data("tea")
str(tea)
dim(tea)
keep_columns <- c("Tea", "How", "how", "sugar", "where", "lunch")
tea_time <- select(tea, keep_columns)
summary(tea_time)
str(tea_time)
gather(tea_time) %>% ggplot(aes(value)) + facet_wrap("key", scales = "free") + geom_bar() + theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8))
```

```{r}
mca <- MCA(tea_time, graph = FALSE)
summary(mca)
plot(mca, invisible=c("ind"), habillage = "quali")
```



# The MCA is a very useful tool to analyze a non-numerical, nominal cathegorcal, qualitative data. Among others, it provides insights into existing patterns in the data. From the example of the MCA above,that explors tea-drinking habits, we can get insights into various aspects of the data at once. Also, some interesting relationships are revealed. For example, we can see that Earl Grey is more likely to be drunk with milk, than lemon or alone. It's also more likely to be drunk with sugar than without. The black tea, on the other hand, is more likely to be taken without sugar, as well as more likely to be enjoyed with lemon than with milk. Most individuals don't drink tea with lunch. 

```{r}
plot(mca, invisible=c("var"), habillage = "quali")
```
# The MCA above shows exclusivly the distribution of data concerning individuals, suggesting similarities and dissimilarities among them. They are arranged in clear clusters. 

```{r}
library(factoextra)
data("tea_time")
res.mca <- MCA(tea_time, graph=FALSE)
fviz_mca_biplot(res.mca, repel = TRUE, ggtheme = theme_minimal())
```

# This extensive plot shows both, variables and individuals at the same time, highliting relationships among them. The distance measures the similarity and dissimilarity among the variables and individuas, that is the possible correlations between them.  




